---
title: "Clone survival screens"
author: "Kimon Froussios"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
params:
  counts: './null'       # provide valid counts file with all the samples.
  summaries: './null'    # provide valid DEseq2-style covariates table file with sample, condition, colour. Can be used to subset the samples included in counts.
  hammdist: '`r NULL`'     # provide valid tidy table for the hamming distance histograms
  outpref: './null'     # provide a destination for the output table
  samples: '`r NULL`'    # provide valid file listing samples, conditions and assigned colours.
  topdf: False           # redirect plot output to PDF file
  refsamps: 1            # Which samples to use as reference abundances for naming the barcodes.
  count_thresh: 50       # Barcodes with counts below this will be discarded.
  abund_thresh: 0.01     # Barcodes exceeding this proportional representation are considered top hits.
  extra_bc: '`r NULL`'
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

Several plots in this report are interactive, so be sure to hover your pointer over them!

# Preparation

Input

```{r libraries, result="hide", warning=FALSE}
library(data.table)
library(magrittr)
library(dplyr) 
library(ggplot2)
library(plotly)
library(ggrepel)
library(RColorBrewer)
library(pheatmap)
library(patchwork)

# params <- list(counts = opt$countsFile, summaries = opt$summariesFile, hammdist=opt$hammingFile, samples = opt$samples, outpref = file.path(opt$resultsDir, sub('.txt|.tsv', '', basename(opt$countsFile))), topdf = opt$topdf, refsamps=opt$refsamps, abund_thresh=opt$abund_thresh, count_thresh=opt$count_thresh, extra_bc= opt$extrabc )

# ggplot2 theme.
gth <- theme(axis.line.x= element_line(),
	         axis.line.y= element_line(),
	         strip.background= element_rect(fill= "grey95"),
	         strip.text.y= element_text(size= rel(1.2)),
	         strip.text.x= element_text(size= rel(1.1)),
	         panel.background= element_rect(fill = "white"),
	         panel.grid.major = element_line(colour="grey95"),
	         panel.grid.minor= element_blank(),
	         legend.key = element_rect(fill = 'white'),
	         plot.margin = unit(c(0.3, 1, 0.3, 0.3),"cm"),
	         axis.text.x = element_text(angle=90, hjust=1),
	         legend.position = "none"
)

getPalette <- colorRampPalette(brewer.pal(8, "Paired"))

# Data
counts <- fread(params$counts)
summaries <- fread(params$summaries)
if (!is.null(params$hammdist)) {
  hammdist <- fread(params$hammdist)
}

# Output
statsout <- params$statsout
topdf <- params$topdf

# Set sample order.
samples_tidy <- fread(params$samples)
setnames(samples_tidy, c("name", "treatment", "colour"))
samples_tidy[, name := factor(name, ordered=TRUE, levels=name)]

# Set treatment colours
sampleNames <- as.character(samples_tidy$name)
tc <- unique(samples_tidy[, .(treatment, colour)])
treatCols <- tc$colour
names(treatCols) <- tc$treatment
samples_tidy[, colour := NULL]
treatCols <- c(treatCols, bc='white')

# If summaries only has one column of values (ie. one BAM file, or already in long format)
if (length(summaries) == 2) {
    names(summaries) <- c('row_ID', 'file')
} else {
  names(summaries)[1] <- 'row_ID' # need a stable name I can reference
  summaries <- melt(summaries, id.vars = "row_ID", variable.name = "file", value.name = "reads")
}

# Plug any gaps from undetected barcodes.
counts[is.na(counts)] <- 0
summaries[is.na(summaries)] <- 0

# Reorder/select samples.
counts <- counts[, c('barcode', sampleNames), with=FALSE]
if (length(summaries) > 2) {
  summaries <- summaries[row_ID %in% c("BCUnmatched", "EmptyVector", "SampleUnknown", "SpikeIn", sampleNames)]
}
if (!is.null(params$hammdist)) {
  hammdist <- hammdist[Sample %in% sampleNames, ]
  hammdist[, Sample := ordered(Sample, levels=sampleNames)]
}

# Assign a numeric ID to barcodes by abundance in the designated samples. +1 is to skip the first column which is barcodes.
setorderv(counts, names(counts)[c(params$refsamps)+1], order=-1)
counts[, bc_id := 1:nrow(counts)]
setkey(counts, bc_id)

# Get rid of barcodes not seen in the samples (as a result of subsetting the samples).
# Do this after assigning bc_ids so bc_id  is consistent across different analyses with the same reference samples.
counts <- counts[rowSums(counts[, 2:length(counts)]) != 0, ]

# Vector of counts columns.
cntCols <- 2:(length(counts)-1)

# Thresholds
count_thresh = params$count_thresh
abund_thresh <- params$abund_thresh
```

Calculate proportions and tidy up the data format.

```{r tidydata}
# Number of reads and number of detected barcodes in each sample.
libsizes <- sapply(counts[, cntCols, with=FALSE], sum)
numBarcodes <- sapply(counts[, cntCols, with=FALSE], function (x) { sum(x>0) } )
numGoodBarcodes <- sapply(counts[, cntCols, with=FALSE], function (x) { sum(x>=count_thresh) } )

# Barcode abundances as fractions of the library sizes
relative <- as.data.table(lapply(names(counts[, -1, with=FALSE]), function(x) { 
                                  if (x=='bc_id') { 
                                    return(counts$bc_id)
                                  } else {          
                                    return(counts[[x]] / libsizes[x])
                                  }} ))
names(relative) <- names(counts[, -1, with=FALSE])
setkey(relative, bc_id)

# Minimum barcode representation threshold
numBarcodesThresh <- sapply(relative[, 1:(length(relative)-1)], function (x) { sum(x >= abund_thresh) } )

# Make tidy summary of counts.
count_summary <- data.table(Count = as.double(c(libsizes, numBarcodes, numGoodBarcodes, numBarcodesThresh)),
                            Type = factor(c(rep('Reads', length(libsizes)), 
                                            rep('all_BCs', length(numBarcodes)),
                                            rep('fltr_BCs', length(numGoodBarcodes)),
                                            rep('top_Barcodes', length(numBarcodesThresh)) ),
                                          levels= c('all_BCs', 'fltr_BCs', 'top_Barcodes', 'Reads')),
                            Sample = c(names(libsizes), names(numBarcodes), names(numGoodBarcodes), names(numBarcodesThresh)) )
count_summary[, Sample := factor(count_summary$Sample, levels=sampleNames, ordered=TRUE)]
setorder(count_summary, Sample)

# Make the proportions tidy too.
relative_tidy <- melt(relative, id.vars='bc_id', variable.name="Sample", value.name="Proportion")
relative_tidy[, Sample := factor(relative_tidy$Sample, levels=sampleNames, ordered=TRUE)]
setorder(relative_tidy, Sample)

# Count filter base donly on the reference (to get around barcode depletions in endpoint samples for correlation purposes)
counts[, refcount_Good := rowMeans(counts[, params$refsamps + 1, with=FALSE]) >= count_thresh]

# I'll need tidy counts too...
# I'll need tidy counts too...
counts_tidy <- melt(counts[, 2:length(counts)], id.vars=c('bc_id', 'refcount_Good'), variable.name="Sample", value.name="Reads")
counts_tidy[, Sample := factor(counts_tidy$Sample, levels=sampleNames, ordered=TRUE)]
setorder(counts_tidy, Sample)

# Now bring it all together.
all_tidy <- merge(counts_tidy, relative_tidy, by=c("bc_id", "Sample"))
all_tidy[, Sample := factor(all_tidy$Sample, levels=sampleNames, ordered=TRUE)]
setorder(all_tidy, Sample)

# rm(relative_tidy)
# rm(counts_tidy)
```

Create categories.

```{r categories}
# Make selection variable for the counts
counts_bool <- as.data.table(lapply(names(counts[, -1, with=FALSE]), function(x) { 
                                  if (x=='bc_id') { 
                                    return(counts$bc_id)
                                  } else {          
                                    return(counts[[x]] >= count_thresh)
                                  }} ))
names(counts_bool) <- names(counts[, -1, with=FALSE])
setkey(counts_bool, bc_id)

counts_bool_tidy <- melt(counts_bool, id.vars='bc_id', variable.name="Sample", value.name="count_Good")


counts_bool_tidy[, Sample := factor(counts_bool_tidy$Sample, levels=sampleNames, ordered=TRUE)]
all_tidy <- merge(all_tidy, counts_bool_tidy, by=c("bc_id", "Sample"))
setorder(all_tidy, Sample)

# rm(counts_bool_tidy)
# rm(counts_bool)

# Make selection variable for the proportions
all_tidy[, prop_Top := Proportion > abund_thresh]

# Make selection variables for treatment mouse and essay.
all_tidy <- merge(all_tidy, samples_tidy, by.x="Sample", by.y="name")
all_tidy[, Sample := factor(all_tidy$Sample, levels=sampleNames, ordered=TRUE)]
setorder(all_tidy, Sample)
# rm(samples_tidy)

# all_tidy has the counts, don't need them here anymore. But keep it for the sequences.
counts <- counts[, .(bc_id, barcode)]
```

```{r table}
fwrite(merge(counts[,.(barcode, bc_id)], 
             dcast(all_tidy, bc_id ~ Sample, value.var = c("Reads", "Proportion")), on="bc_id"),
       file=paste0(params$outpref, '_data.tsv'), sep="\t", quote=FALSE)
```

```{r pdfstart}

if (topdf){
  pdf(paste0(params$outpref, '_report.pdf'), paper="a4", width=7, height=10)
}
```

# Overview of the number of reads allocated per sample

```{r, fig.height=6, fig.width=7}
summaries <- melt(summaries, id.vars = "row_ID", variable.name = "file", value.name = "reads")
summaries[, total := sum(.SD$reads), by=file]
summaries[, rate := reads / total]
names(summaries)[1] <- "sample"
summaries <- summaries[reads > 0,]
# summaries$file <- ordered(summaries$file, labels=sampleNames)

clr <- getPalette(length(unique(summaries$sample)))
names(clr) <- unique(summaries$sample)
clr[c("BCUnmatched", "EmptyVector")] <- c("black", "white")

p1 <- ggplot(summaries) +
    geom_bar(aes(x=file, y=rate, fill=sample), stat="identity", position="stack", colour='black', size=0.25) +
    scale_fill_manual(values=clr) +
    labs(title="Efficiency of barcode detection and sample demultiplication", x='File', y='Fraction of reads') +
    gth + theme(legend.position = "bottom")

if (topdf){
  print ( p1 )
} else {
  ggplotly( p1, dynamicTicks = TRUE)
}
```

# Number of reads and barcodes

```{r}
fwrite(dcast(count_summary[(Type != "top_Barcodes"), ], Sample ~ Type, value.var = 'Count'),
       file=paste0(params$outpref, '_summary.tsv'), sep="\t", quote=FALSE)
```

```{r counts, fig.height=6, fig.width=7}
# Basic faceted barplot
p1 <- ggplot(data=droplevels(count_summary[(Type != "top_Barcodes"), ]), aes(x=Sample, y=Count, fill=Type))  +
            facet_grid(Type ~ ., scales = "free") +
            geom_bar(stat='identity', width=0.7, size=0.2) +
            scale_fill_manual(values = c(all_BCs = "mediumorchid2", 
                                         fltr_BCs = "mediumorchid3", 
                                         Reads= "darkgoldenrod2")) +
            scale_y_continuous(expand = expand_scale(mult = c(0, .15)), limits = c(0, NA)) +
            labs(title="Total # of reads, and # of detected barcodes", x='', y='') +
            gth + theme(legend.position = "none")

if (topdf){
  print ( p1 )
} else {
  ggplotly( p1 )
}
```

```{r counts2, fig.height=6, fig.width=7}
# Same but sqrt-scaled
if (topdf){
  p1 <- ggplot(data=droplevels(count_summary[(Type != "top_Barcodes"), ]), aes(x=Sample, y=Count, fill=Type))  +
          facet_grid(Type ~ ., scales = "free") +
          geom_bar(stat='identity', width=0.7, size=0.2) +
          scale_fill_manual(values = c(all_BCs = "mediumorchid2", 
                                       fltr_BCs = "mediumorchid3", 
                                       Reads= "darkgoldenrod2")) +
          scale_y_continuous(expand = expand_scale(mult = c(0, .15)), limits = c(0, NA), trans='sqrt') +
          labs(title="Total # of reads, and # of detected barcodes", x='', y='') +
          gth + theme(legend.position = "none")
  print ( p1 )
} 
```

# Correlations 

Correlate using only the barcodes that are on average above threshold in the reference sample(s).

```{r}
cf <- dcast(all_tidy[(refcount_Good), ], bc_id ~ Sample, value.var='Reads', fill=NA_integer_)

cat(paste('Number of barcodes used:', nrow(cf)))

print( pheatmap( cor(cf[, 2:ncol(cf)]), cluster_rows = FALSE, cluster_cols = FALSE) )
```

# Distribution of barcode abundances above threshold

Mind that the Y-axis scale is **logarithmic**, meaning the distributions are much more top-skewed than they appear.

```{r proportions}
p1 <- all_tidy %>% 
    filter(count_Good) %>%
    ggplot(aes(x=Sample, y=Proportion)) +
      geom_violin(aes(fill=treatment), width=1) +
      geom_boxplot(fill="white", width=0.1, outlier.colour="transparent") +
      geom_hline(yintercept = abund_thresh, linetype='dotted') +
      scale_y_log10() +
			annotation_logticks(base=10, sides='lr') +
      scale_fill_manual(values=treatCols) +
      labs(title="Barcode abundances", subtitle=paste("for barcodes with at least", count_thresh, "reads"), x='') +
  		gth

# if (topdf){
  print ( p1 )
# } else {
#   ggplotly( p1, dynamicTicks = TRUE )
# }
```

# Barcode Complexity

```{r complexity}
# Identify shared top barcodes
sharedBarcodes <- all_tidy %>%
  filter(prop_Top) %>%
  group_by(bc_id) %>%
  summarise(numSamples=n()) %>%
  filter(numSamples > 1)

# Create selection variable for shared top-ness.
all_tidy[, is_Shared := bc_id %in% sharedBarcodes$bc_id]
```

Top barcodes that are shared among at least 2 samples.

```{r complexity2}
print( as.data.table(all_tidy %>% 
         filter(is_Shared, prop_Top) %>%
         group_by(bc_id) %>% 
         summarise(Samples = paste(Sample, collapse = " ; ")))
)
```

The colours identify shared barcodes across samples. Grey means the barcode is not shared. 
The barcodes that are below the threshold are represent by the uncoloured top compartment in each stack.

```{r complexity_plots2, fig.height=6}
# Prepare colour fill.
all_tidy[, sharedFill := as.character(ifelse(bc_id %in% sharedBarcodes$bc_id, bc_id, 'notshared'))]

# Numbers of Barcodes
secScale <- max(numBarcodesThresh) / max(numBarcodes)
count_summary[, scaledCount := Count / secScale]

# Filter data, then add dummy entries for any samples that have no qualifying barcodes, so they are still shown on the axis.
# Also fill-in a bulk segment for the barcodes that did not meet the threshold.
tmp <- as.data.table(all_tidy %>% 
  filter(prop_Top) %>%
  arrange(desc(Proportion)))
dropped <- unique(all_tidy$Sample)[! unique(all_tidy$Sample) %in% unique(tmp$Sample)]
tmp <- rbind(tmp, data.frame(Sample=dropped, bc_id=0, refcount_Good=0, Reads=0, Proportion=0, count_Good=NA, prop_Top=FALSE, treatment=NA, is_Shared=NA, sharedFill=NA) )
fillup <- tmp[, .(Sample, Proportion)] [, 1-sum(Proportion), by=Sample]
setnames(fillup, c("Sample", "Proportion"))
tmp <- rbind(tmp, cbind(fillup, data.frame(bc_id=0, refcount_Good=0, Reads=0, count_Good=NA, prop_Top=FALSE, treatment=NA, is_Shared=NA, sharedFill='fillup')) )

# Assign colours
elem <- as.character(unique(tmp$sharedFill[!is.na(tmp$sharedFill) & ! tmp$sharedFill %in% c('notshared', 'fillup')]))
elemcol <- getPalette(length(elem))
names(elemcol) <- elem
elemcol["notshared"] <- "grey70"
elemcol["fillup"] <- "grey95"

p1 <- ggplot(data=tmp, aes(x=Sample, y=Proportion, fill=sharedFill, group="A", label=bc_id)) +
         geom_bar(stat='identity', position=position_stack(), colour="black", width=0.9, size=0.2) +
         scale_y_continuous(expand = c(0, 0), limits= c(0,1),
                            sec.axis = sec_axis(trans = ~ .)) +
         scale_fill_manual(values=elemcol) +
         ylab("Individual barcode proportion") +
         labs(title=paste0("Barcodes with at least ", abund_thresh * 100,"% representation"), x='') +
         gth + theme(panel.grid.major.x = element_blank(), legend.position="none")

if (topdf){
  print ( p1 )
} else {
  ggplotly( p1, dynamicTicks = TRUE )
}
```

Number of barcodes passing the counts threshold vs. number of barcodes passing the representation threshold.

```{r complexity_plots3}
# The labels can create terrible clutter if the plot is small with lots of points. 
# Plotly enables hover-over to get Sample info
if (topdf){
  print(
    ggplot(data=all_tidy %>%
                  filter(count_Good) %>%
                  group_by(Sample) %>%
                  summarise(n_Barcodes=n(), n_TopBarcodes=sum(prop_Top), Treatment=unique(treatment)),
           aes(label=Sample, x=n_Barcodes, y=n_TopBarcodes, fill=Treatment)) +
      geom_label_repel(colour="black", fill="white", size=3, direction="both", force=1, min.segment.length = 0) +
      geom_point(shape=21) +
      scale_fill_manual(values=treatCols) +
      ggtitle("Barcode Diversity") +
      gth
  )
} else {
  ggplotly(
    ggplot(data=all_tidy %>% 
                  filter(count_Good) %>%
                  group_by(Sample) %>% 
                  summarise(n_Barcodes=n(), n_TopBarcodes=sum(prop_Top), Treatment=unique(treatment)), 
           aes(label=Sample, x=n_Barcodes, y=n_TopBarcodes, colour=Treatment)) +
      geom_point(shape=16, size=1, alpha=0.6) +
      scale_colour_manual(values=treatCols) +
      ggtitle("Barcode Diversity") +
      gth,
    dynamicTicks = TRUE
  )
}
```

# Enrichment of barcodes

## Expression heatmap of top barcodes.

All barcodes above the proportion threshold.

```{r enrichment, fig.height=15}
top <- all_tidy %>% 
  filter(prop_Top) %>%
  select(bc_id) %>%
  arrange(bc_id) %>%
  unique()

p1 <- all_tidy %>%
  arrange(bc_id) %>%
  filter(bc_id %in% top$bc_id) %>%
  select(Sample, bc_id, Proportion) %>%
  mutate(bc_id=factor(as.character(bc_id), levels=as.character(top$bc_id))) %>%
  ggplot(aes(x=Sample, y=bc_id, fill=Proportion)) +
    geom_tile(colour='transparent') +
    # scale_fill_gradient(low="transparent", high="midnightblue") +
    labs(x='') +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1),
          legend.position = "right",
          panel.background = element_rect(fill="white"),
          panel.grid.major = element_line(colour="grey95"))

if (topdf){
  print ( p1 )
} else {
  ggplotly( p1, dynamicTicks = TRUE )
}
```

## Distribution of top shared barcodes

Mind that the Y-axis scale is **logarithmic**. Outliers at the top are much farther away from the main bulk of the distribution than they appear.

```{r enrichment_shared}
for (bc in sharedBarcodes$bc_id) {
  # bc <- sharedBarcodes$bc_id[1]
  
  # Create dummy sample with the abundances of the given barcode only.
  # Preserve sample order.
  tmp <- all_tidy[(bc_id==bc),]
  bct <- paste("bc:", bc)
  tmp$Sample <- factor(bct, levels=c(sampleNames, bct), ordered=TRUE)
  tmp$treatment <- "bc"
  tmp2 <- all_tidy[(count_Good),]
  tmp2$Sample <- factor(tmp2$Sample, levels=c(sampleNames, bct), ordered=TRUE)
  tmp <- rbind(tmp2, tmp)
  
  print(paste('Barcode', bc))
  print(counts$barcode[counts$bc_id==bc])
  # print(
  #   ggplot(data=tmp, aes(x=Sample, y=Proportion, fill=treatment)) +
  #     geom_hline(yintercept=abund_thresh, linetype='dotted') +
  #     geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  #     geom_point(data=all_tidy[(bc_id == bc), ], fill=treatCols['bc'], size=2.2, shape=23) +
  #     scale_y_log10() +
  #     scale_fill_manual(values=treatCols) +
  #     ggtitle(paste("Abundance of barcode",bc)) +
  #     gth + theme(panel.grid.minor.y = element_line(colour='grey95'))
  # )
  print(
    ggplot(data=tmp, aes(x=Sample, y=Proportion, fill=treatment)) +
      geom_hline(yintercept=abund_thresh, linetype='dotted') +
      geom_violin() +
      geom_boxplot(fill='white', width=0.15, outlier.size = 0.8, outlier.alpha = 0.5, alpha=0.4) +
      geom_point(data=all_tidy[(bc_id == bc), ], fill=treatCols['bc'], size=2.2, shape=23) +
      scale_y_log10() +
      scale_fill_manual(values=treatCols) +
      labs(paste("Abundance of barcode",bc), x='') +
      gth + theme(panel.grid.minor.y = element_line(colour='grey95'))
  )
  # print(
  #   ggplot(data=tmp, aes(x=Sample, y=Proportion, fill=treatment)) +
  #     geom_hline(yintercept=abund_thresh, linetype='dotted') +
  #     geom_boxplot(outlier.size = 0.8, outlier.alpha = 0.5) +
  #     geom_point(data=all_tidy[(bc_id == bc), ], fill=treatCols['bc'], size=2.2, shape=23) +
  #     scale_y_log10() +
  #     scale_fill_manual(values=treatCols) +
  #     annotation_logticks(base=10, sides='lr') +
  #     labs(paste("Abundance of barcode",bc), x='') +
  #     gth
  # )
}
```

## Distribution of manually selected barcodes

Mind that the Y-axis scale is **logarithmic**. Outliers at the top are much farther away from the main bulk of the distribution than they appear.

```{r enrichment_manual}
if (is.null(params$extra_bc)) {
  cat("No additional barcodes selected")
} else {
	for (bc in params$extra_bc) {
	  # bc <- sharedBarcodes$bc_id[1]
	  
	  # Create dummy sample with the abundances of the given barcode only.
	  # Preserve sample order.
	  tmp <- all_tidy[(bc_id==bc),]
	  bct <- paste("bc:", bc)
	  tmp$Sample <- factor(bct, levels=c(sampleNames, bct), ordered=TRUE)
	  tmp$treatment <- "bc"
	  tmp2 <- all_tidy[(count_Good),]
	  tmp2$Sample <- factor(tmp2$Sample, levels=c(sampleNames, bct), ordered=TRUE)
	  tmp <- rbind(tmp2, tmp)
	  
	  print(paste('Barcode', bc))
	  print(counts$barcode[counts$bc_id==bc])
	  # print(
	  #   ggplot(data=tmp, aes(x=Sample, y=Proportion, fill=treatment)) +
	  #     geom_hline(yintercept=abund_thresh, linetype='dotted') +
	  #     geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
	  #     geom_point(data=all_tidy[(bc_id == bc), ], fill=treatCols['bc'], size=2.2, shape=23) +
	  #     scale_y_log10() +
	  #     scale_fill_manual(values=treatCols) +
	  #     ggtitle(paste("Abundance of barcode",bc)) +
	  #     gth + theme(panel.grid.minor.y = element_line(colour='grey95'))
	  # )
	  print(
	    ggplot(data=tmp, aes(x=Sample, y=Proportion, fill=treatment)) +
	      geom_hline(yintercept=abund_thresh, linetype='dotted') +
	      geom_violin() +
	      geom_boxplot(fill='white', width=0.15, outlier.size = 0.8, outlier.alpha = 0.5, alpha=0.4) +
	      geom_point(data=all_tidy[(bc_id == bc), ], fill=treatCols['bc'], size=2.2, shape=23) +
	      scale_y_log10() +
	      scale_fill_manual(values=treatCols) +
	      labs(paste("Abundance of barcode",bc), x='') +
	      gth + theme(panel.grid.minor.y = element_line(colour='grey95'))
	  )
	  # print(
	  #   ggplot(data=tmp, aes(x=Sample, y=Proportion, fill=treatment)) +
	  #     geom_hline(yintercept=abund_thresh, linetype='dotted') +
	  #     geom_boxplot(outlier.size = 0.8, outlier.alpha = 0.5) +
	  #     geom_point(data=all_tidy[(bc_id == bc), ], fill=treatCols['bc'], size=2.2, shape=23) +
	  #     scale_y_log10() +
	  #     scale_fill_manual(values=treatCols) +
	  #     annotation_logticks(base=10, sides='lr') +
	  #     labs(paste("Abundance of barcode",bc), x='') +
	  #     gth
	  # )
	}
}
```

# Barcode sequence similarity

If merging of the barcodes by Hamming distance was enabled during processing, the frequency of
pair-wise barcode similarity levels, as measured by the edit distance, will be shown in this
section. Each barcode pair is counted once, and all the possible pairs are counted. The values
tracked are the number of barcode pairs in each distance bin and the respective numbers of reads.

```{r}
if (is.null(params$hammdist)) {
  cat("No info available")
} else {
  library(ggridges)
  
  hammdist[, clr := vapply(as.character(hammdist$Sample), function(x) { samples_tidy[name==x, colour] }, character(1)) ]
  
  print( ggplot(hammdist, aes(x=HammDist, y=Sample, height=sqrt(BCs), colour=clr, fill=clr) ) +
    geom_density_ridges(stat="identity", scale=1, alpha=0.5) +
    scale_colour_identity() +
    scale_fill_identity() +
    labs(title="Distribution of pairwise barcode distances", x = "Hamming distance") +
    gth + theme(axis.text.x = element_text(angle=0))
  )
  
  p1 <- ggplot(hammdist, aes(x=HammDist, group=Sample, y=BCs, colour=clr) ) +
    geom_line() +
    scale_colour_identity() + 
    scale_fill_identity() +
    labs(x = "Hamming distance", y="Number of barcode pairs") +
    gth + theme(axis.text.x = element_text(angle=0))
  p2 <- p1 +
    scale_y_log10() +
    annotation_logticks(base=10, sides='lr') +
    labs(y='')
  print(p1 + p2)
}
```

# Session Info

```{r}
sessionInfo()
```


```{r pdfend}
if (topdf) {
  dev.off()
}
```
